---
title: "Hurricane Report"
author: "Jerry Chen and Jocelyn Hunyadi"
date: "12/4/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}
# Acquire Necessary R Packages:
require(stringr)
require(mdsr)
require(tidyverse)
require(mosaic)
require(tidyr)
require(dplyr)
require(readxl)
require(rvest)
require(lubridate)

```

# Abstract

> Summary inserted below:
This project aims to analyze the paths of hurricanes that made landfall on the Continental United States over the past decade in the context of damage and deaths. More emphasis will be placed on the most significant and intense hurricanes including Hurricane Katrina and Sandy. We hope to illustrate these hurricanes dynamically via animations that highlight both the paths of the hurricanes as well as the severity of the damage and number of fatalities caused. Additionally, we hope to analyze trends in both frequency and intensity of hurricanes in the context of climate change with surface water temperature and CO2 emissions as possible explanatory variables in a predictive model of hurricane frequency.

> one paragraph
> Summarize Goals/Results


# Introduction

> What is/are the research question(s)
> Why is this interesting/important?
> What contributions have we made to answer this question?
  > Overview of the Model

# Data

> Brief Description of dataset(s)
  > Variables?
  > Where did the data come from?
  > Units for each measurment
  > Population that was sampled
> Read in the Data and add new variables
> Univariate Analysis

```{r, echo = FALSE}
# Obtain Hurricane Path Data from NOAA Hurricane Reports:

```

```{r, echo = FALSE}
# Obtain CO2 Data - Found originally in txt file:
CO2_data <- read.csv("CO2_data.csv")
```

```{r, echo = FALSE}
# Scrape the Data for the Costliest Atlantic Hurricanes from Wikipedia:
url <- 'https://en.wikipedia.org/wiki/List_of_costliest_Atlantic_hurricanes'
list <- url %>%
  read_html() %>%
  html_nodes("table")
  
  # Select the correct table to a dataframe:
costly_hurr <- html_table(list[[1]], fill = TRUE)

  # Join with CO2 Data:
costly_hurr <- left_join(costly_hurr, CO2_data, by = c("Season" = "Year"))
costly_hurr <- costly_hurr[,c(1:6,8)]

  # Modify column names:
colnames(costly_hurr) = c("name","damage","year","classification","area","reference","co2")

# Save Harvey Damage:
actual_Harvey <- readr::parse_number(substring(costly_hurr$damage[1],2))

# Modifications to the Data:
  # Remove entries with NA in co2 column:
costly_hurr <- costly_hurr %>% filter(!is.na(co2)) %>%
  # Extract numbers from damage column and convert to numeric:
  mutate(damage = readr::parse_number(substring(damage,2)))

  # Remove part of string excess from classification column:
costly_hurr <- costly_hurr %>%
  mutate(classification = substring(classification,12))

  # Add Category variable
costly_hurr <- costly_hurr %>%
  mutate(category = ifelse(classification == "storm", 0, substring(classification,2,2))) %>%
  mutate(category = ifelse(category == "s", 0, readr::parse_number(category))) %>%
  mutate(category_factor = as.factor(category))

```

```{r, echo = FALSE}
# Scrape the Data for each Atlantic Hurricane Season since 1910:
  # Data originally obtained from the following url:
  # https://en.wikipedia.org/wiki/Atlantic_hurricane_season#Number_of_tropical_storms_and_hurricanes_per_season
  # Data was Modified - modified data = Hurricane_Seasons.csv
  # Please See the Appendix to see how data was modified
seasons_hurr <- read_csv("Hurricane_Seasons.csv")
seasons_hurr <- seasons_hurr[,2:10]

```


# Results

> What does model tell you about the research question?
> Interpret coefficients in context/explain relevance

```{r}
# Looking at AVERAGE damage for hurricanes that occurred in the same year:
costly_hurr_2 <- costly_hurr %>%
  group_by(co2) %>%
  summarise(avg_damage = mean(damage), total_damage = sum(damage), max_damage = max(damage)) %>%
  mutate(sqrt_avg_damage = sqrt(avg_damage), sqrt_max_damage = sqrt(max_damage))

mod_avg <- lm(sqrt_avg_damage ~ co2, data = costly_hurr_2)
summary(mod_avg)

ggplot(costly_hurr_2, aes(y = sqrt_avg_damage, x = co2)) + geom_point() + labs(title = "Average Damage by Costly Hurricanes", x = "CO2 (ppm)", y = "Square Root Average Damage (Billions)")
```

```{r}
# Category - no indicators were significant, 
# Is major:
costly_hurr <- costly_hurr %>% mutate(is_major = ifelse(category >= 3, TRUE, FALSE),
                                sqrt_damage = sqrt(damage))

mod_h2 <- lm(sqrt_damage ~ is_major, data = costly_hurr)
anova(mod_h2) # at .1 for sqrt_damage.
```

```{r}
# Looking at MAX damage for costly hurricanes:
mod_max <- lm(sqrt_max_damage ~ co2, data = costly_hurr_2)
summary(mod_max)

ggplot(costly_hurr_2, aes(y = sqrt_max_damage, x = co2)) + geom_point() + labs(title = "Damage by the Most Costly Hurricane", x = "CO2 (ppm)", y = "Sqare Root Max Damage (Billions)")
```


# Diagnostics

> Full diagnostics of the model
> Predictions

```{r}
# Predicting Harvey Damage from Average:
newdata <- data.frame(co2 = 403.38)
  # Must square - y = sqrt_damage:
p1 <- (predict(mod_avg, newdata, interval = "predict"))^2
  # Difference with Upper:
avg_upper_diff <- actual_Harvey - p1[3]; avg_upper_diff # off by 152 billion
  # Difference with Lower:
avg_lower_diff <- actual_Harvey - p1[2]; avg_lower_diff # off by 198.3037 billion
  # Not Very Useful. Also, doesn't quite make as much sense to use the Average.

```

```{r}
# Predicting Harvey Damage from Maximum:
p2 <-(predict(mod_max, newdata, interval = "predict"))^2
  # Difference with Upper:
max_upper_diff <- actual_Harvey - p2[3]; max_upper_diff # off by 94.6 billion
  # Difference with Lower:
max_lower_diff <- actual_Harvey - p2[2]; max_lower_diff # off by 198 billion
  # Actual Fit:
max_actual_diff <- actual_Harvey - p2[1]; max_actual_diff # off by 173.28 billion
  # Better, but still off by a lot.
  
  # Lurking Variables may account for unaccounted for variation in damage.
```

# Conclusion

> First - remind of question and provide summary of findings
> Second - Discussion of model limitations
  > What can be done to improve the model?
  > What can be done to improve the data?

# Appendix

### Modifications to Hurricane Season Data:
```{r}
url <- "https://en.wikipedia.org/wiki/Atlantic_hurricane_season#Number_of_tropical_storms_and_hurricanes_per_season"

list <- url %>%
  read_html() %>%
  html_nodes("table")
  
# Get Data:
names <- c("year","num_trop_storms","num_hurricanes","num_major","deaths","damage")

h_season01 <- html_table(list[[8]], fill = TRUE)
h_season01 <- h_season01[,1:6]
colnames(h_season01) <- names

h_season02 <- html_table(list[[9]], fill = TRUE)
h_season02 <- h_season02[,1:6]
colnames(h_season02) <- names

h_season03 <- html_table(list[[10]], fill = TRUE)
h_season03 <- h_season03[,1:6]
colnames(h_season03) <- names

h_season04 <- html_table(list[[11]], fill = TRUE)
h_season04 <- h_season04[,1:6]
colnames(h_season04) <- names

h_season05 <- html_table(list[[12]], fill = TRUE)
h_season05 <- h_season05[,1:6]
colnames(h_season05) <- names

h_season06 <- html_table(list[[13]], fill = TRUE)
h_season06 <- h_season06[,1:6]
colnames(h_season06) <- names

h_season07 <- html_table(list[[14]], fill = TRUE)
h_season07 <- h_season07[,1:6]
colnames(h_season07) <- names

h_season08 <- html_table(list[[15]], fill = TRUE)
h_season08 <- h_season08[,c(1,3:7)]
colnames(h_season08) <- names

h_season09 <- html_table(list[[16]], fill = TRUE)
h_season09 <- h_season09[,c(1,3:7)]
colnames(h_season09) <- names

h_season10 <- html_table(list[[17]], fill = TRUE)
h_season10 <- h_season10[,c(1,3:7)]
colnames(h_season10) <- names

h_season11 <- html_table(list[[18]], fill = TRUE)
h_season11 <- h_season11[,c(1,3:7)]
colnames(h_season11) <- names

# Join together:
h_season <- rbind(h_season01, h_season02, h_season03, h_season04, h_season05, h_season06, h_season07, h_season08, h_season09, h_season10, h_season11)

# Remove entries with Unknown damage:
h_season <- h_season %>% filter(damage != "Unknown")

# Remove symbols/words from Damage and multiply by appropriate multiplier column:
h_season <- h_season %>%
  mutate(damage_number = readr::parse_number(str_extract(damage, "[0-9|\\.]+")),
         damage_magnitude = str_extract(damage, "[m|b]illion"),
         damage_magnitude = ifelse(is.na(damage_magnitude), "thousand", damage_magnitude),
         damage = ifelse(damage_magnitude == "thousand", damage_number*1000, 
                         ifelse(damage_magnitude == "million", damage_number*1000000, damage_number*1000000000)))

# Add total storms column:
h_season <- h_season %>% mutate(storms_total = num_trop_storms + num_hurricanes)
```


